---
title: "Getting Started with tradefeatures"
author: "TradingVerse Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with tradefeatures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

Welcome to `tradefeatures`! This package provides comprehensive technical analysis indicators for quantitative trading in R. Whether you're a beginner learning technical analysis or an experienced trader building sophisticated strategies, tradefeatures makes it easy to add indicators to your market data.

## Installation

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("tradingverse/tradefeatures")

# Load the package
library(tradefeatures)
```

## Quick Example

Here's a complete example showing how to fetch data and add multiple indicators:

```{r eval=FALSE}
library(tradeio)
library(tradefeatures)
library(dplyr)

# Fetch Apple stock data
aapl <- fetch_prices("AAPL", from = "2024-01-01") |>
  add_sma(20) |>
  add_rsi(14) |>
  add_macd() |>
  add_bbands(20)

# View enriched data
head(aapl)
```

## Core Concepts

### Pipe-Friendly Design

All `tradefeatures` functions work seamlessly with pipes (`|>` or `%>%`):

```{r eval=FALSE}
enriched_data <- raw_data |>
  add_sma(20) |>
  add_ema(50) |>
  add_rsi(14)
```

### Automatic Column Management

Indicators are added as new columns to your data:

```{r eval=FALSE}
# Before
names(prices)
# [1] "symbol" "datetime" "open" "high" "low" "close" "volume"

# After adding indicators
prices <- prices |> add_sma(20) |> add_rsi(14)
names(prices)
# [1] "symbol" "datetime" "open" "high" "low" "close" "volume" "sma_20" "rsi"
```

### Multi-Symbol Support

All indicators automatically handle multiple symbols:

```{r eval=FALSE}
# Fetch multiple stocks
stocks <- fetch_prices(c("AAPL", "MSFT", "GOOGL"), from = "2024-01-01")

# Indicators calculated separately for each symbol
stocks <- stocks |>
  add_sma(20) |>
  add_rsi(14)
```

## Indicator Categories

### 1. Trend Indicators

Trend indicators help identify the direction and strength of price movements.

#### Simple Moving Average (SMA)

```{r eval=FALSE}
# Add 20-day SMA
prices |> add_sma(20)

# Multiple SMAs with custom names
prices |>
  add_sma(20, name = "sma_short") |>
  add_sma(50, name = "sma_medium") |>
  add_sma(200, name = "sma_long")
```

#### Exponential Moving Average (EMA)

```{r eval=FALSE}
# Add 12-day EMA
prices |> add_ema(12)

# EMA is more responsive to recent prices
prices |>
  add_ema(12, name = "ema_fast") |>
  add_ema(26, name = "ema_slow")
```

#### Volume Weighted Average Price (VWAP)

```{r eval=FALSE}
# Add VWAP
prices |> add_vwap()
```

### 2. Momentum Indicators

Momentum indicators measure the rate of price change.

#### Relative Strength Index (RSI)

```{r eval=FALSE}
# Add 14-period RSI
prices |> add_rsi(14)

# Identify overbought/oversold
prices |>
  add_rsi(14) |>
  mutate(
    oversold = rsi < 30,
    overbought = rsi > 70
  )
```

#### MACD

```{r eval=FALSE}
# Add MACD with default parameters (12, 26, 9)
prices |> add_macd()

# Custom parameters
prices |> add_macd(fast = 8, slow = 21, signal = 5)

# Detect crossovers
prices |>
  add_macd() |>
  mutate(
    bullish = (macd > macd_signal) & (lag(macd) <= lag(macd_signal))
  )
```

#### Stochastic Oscillator

```{r eval=FALSE}
# Add Stochastic
prices |> add_stochastic(14, smooth = 3)
```

### 3. Volatility Indicators

Volatility indicators measure price variability.

#### Bollinger Bands

```{r eval=FALSE}
# Add Bollinger Bands (20-day, 2 standard deviations)
prices |> add_bbands(20, sd = 2)

# Identify when price touches bands
prices |>
  add_bbands(20) |>
  mutate(
    at_lower = close <= bb_lower,
    at_upper = close >= bb_upper
  )
```

#### Average True Range (ATR)

```{r eval=FALSE}
# Add 14-period ATR
prices |> add_atr(14)

# Use for position sizing
prices |>
  add_atr(14) |>
  mutate(
    position_size = 1000 / (2 * atr)  # Risk 2 ATRs
  )
```

### 4. Volume Indicators

#### On-Balance Volume (OBV)

```{r eval=FALSE}
# Add OBV
prices |> add_obv()

# Look for divergences
prices |>
  add_obv() |>
  mutate(
    price_up = close > lag(close, 10),
    obv_up = obv > lag(obv, 10),
    bullish_divergence = !price_up & obv_up
  )
```

## Pattern Detection

Identify chart patterns automatically:

```{r eval=FALSE}
# Detect moving average crossovers
prices |>
  add_sma(50) |>
  add_sma(200) |>
  mutate(
    golden_cross = detect_golden_cross(sma_50, sma_200),
    death_cross = detect_death_cross(sma_50, sma_200)
  )
```

## Complete Example: Multi-Indicator Analysis

Here's a comprehensive example combining multiple indicators:

```{r eval=FALSE}
library(tradeio)
library(tradefeatures)
library(tradeengine)
library(dplyr)

# Fetch and enrich data
analysis <- fetch_prices("AAPL", from = "2023-01-01") |>
  # Trend
  add_sma(20) |>
  add_ema(50) |>
  # Momentum
  add_rsi(14) |>
  add_macd() |>
  # Volatility
  add_bbands(20) |>
  add_atr(14) |>
  # Volume
  add_obv() |>
  # Create trading signals
  mutate(
    # Trend signal
    uptrend = close > sma_20 & sma_20 > ema_50,
    # Momentum signal
    momentum_up = rsi > 50 & macd > macd_signal,
    # Not overbought
    not_overbought = rsi < 70,
    # Combined signal
    buy_signal = uptrend & momentum_up & not_overbought
  )

# Count signals
sum(analysis$buy_signal, na.rm = TRUE)
```

## Integration with tradeengine

Use indicators in backtesting strategies:

```{r eval=FALSE}
# Complete trading strategy
results <- fetch_prices("AAPL", from = "2023-01-01") |>
  add_sma(20) |>
  add_rsi(14) |>
  add_strategy(
    entry = (close > sma_20) & (rsi < 30),
    exit = (close < sma_20) | (rsi > 70)
  ) |>
  backtest(initial_capital = 10000)

print(results$summary)
```

## Best Practices

### 1. Start Simple

Begin with one or two indicators and understand them well:

```{r eval=FALSE}
# Simple SMA crossover
prices |>
  add_sma(20) |>
  add_sma(50) |>
  mutate(signal = sma_20 > sma_50)
```

### 2. Combine Complementary Indicators

Use indicators from different categories:

```{r eval=FALSE}
# Trend + Momentum + Volatility
prices |>
  add_sma(20) |>        # Trend
  add_rsi(14) |>        # Momentum
  add_atr(14)           # Volatility
```

### 3. Validate with Backtesting

Always test indicator combinations:

```{r eval=FALSE}
results <- prices |>
  add_rsi(14) |>
  add_strategy(entry = rsi < 30, exit = rsi > 70) |>
  backtest(initial_capital = 10000)
```

### 4. Watch for Overfitting

Too many indicators can lead to overfitting:

```{r eval=FALSE}
# Good: 3-4 complementary indicators
prices |> add_sma(20) |> add_rsi(14) |> add_atr(14)

# Risky: Too many similar indicators
prices |> add_sma(10) |> add_sma(20) |> add_sma(30) |> add_sma(50) |> # ...
```

## Next Steps

Explore more advanced topics:

- `vignette("case-study-sma-crossover")` - Complete SMA strategy
- `vignette("case-study-rsi-mean-reversion")` - RSI trading strategy
- `vignette("case-study-multi-indicator")` - Advanced combinations
- `vignette("momentum-indicators")` - Deep dive into momentum
- `vignette("volatility-analysis")` - Understanding volatility

## Summary

`tradefeatures` makes technical analysis in R:

✅ **Easy** - Simple, consistent API  
✅ **Fast** - Optimized calculations  
✅ **Flexible** - 30+ indicators  
✅ **Integrated** - Works with tradeio and tradeengine  
✅ **Documented** - Comprehensive examples  

Happy trading! 📈
